---
title: "Predicción desembolsos"
author: "Showdown Analytics"
date: "March 13, 2021"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Librerías

```{r}
library(tidymodels)
library(modeltime)
library(tidyverse)
library(timetk)
library(lubridate)
```

# Directorio activo

```{r}
setwd("E:/github/RCI-desembolsos/binaries")
```


# Lectura de datos

```{r}
desembolsos_raw <- data.table::fread("G:/SASData/Andres_Franco/AA_Insumos/Mina_Marketing/RCI_DESEMBOLSOS_01.csv") %>% 
  as_tibble()
```

# ETL

```{r}
desembolsos <- desembolsos_raw %>% 
  mutate(fecha_desembolso = dmy(`Fecha Desembolso`),
         zona_id = case_when(is.na(`Zona RCI`) ~ "desconocida",
                                          TRUE ~ as.character(`Zona RCI`)),
         estado_id = `N/U`)
```


# Definir time series semanal

```{r}
desembolsos %>% glimpse()

desembolsos_ts <- desembolsos %>%
  filter(!is.na(fecha_desembolso)) %>% 
  group_by(fecha_desembolso, zona_id, estado_id) %>%
  summarize(n_desem = n()) %>% 
  ungroup()
```

# Timetk diagnostics

```{r}
# Time series
desembolsos_ts %>%
  filter(zona_id == "1") %>% 
  plot_time_series(fecha_desembolso, n_desem)

# Anomaly diagnostics
desembolsos_ts %>%
  filter(zona_id == "1") %>% 
  plot_anomaly_diagnostics(fecha_desembolso, n_desem)

# Seasonal diagnostics
desembolsos_ts %>%
  filter(zona_id == "1") %>% 
  plot_seasonal_diagnostics(fecha_desembolso, n_desem)

# ACF y PACF
desembolsos_ts %>%
  filter(zona_id == "1") %>% 
  plot_acf_diagnostics(fecha_desembolso, n_desem)
```

# Prepare full data 

```{r}
#Perform hierarchical aggregations
full_data_tbl <- desembolsos_ts %>% 
  add_column(all_zones_id = "all_zones", .before = 2) %>% 
  pivot_longer(cols = ends_with("_id"),
               names_to = "category",
               values_to = "identifier") %>% 
  group_by(category, identifier, fecha_desembolso) %>% 
  summarize(n_desem = sum(n_desem, na.rm = TRUE)) %>% 
  ungroup() %>%
  
  ############## FEATURE ENGINEERING
  group_by(category, identifier) %>% 
  
  # FILL MISSING DAYS WITH 0
  pad_by_time(fecha_desembolso, .by = "day", .pad_value = 0) %>%
  
  # CREATE A FUTURE FRAME TO MAKE A MONTHLY PREDICTION
  future_frame(fecha_desembolso, .length_out = 28, .bind_data = TRUE) %>%
  
  # CREATE LAGS FOR THE DESEM
  tk_augment_lags(n_desem, .lags = 28) %>%
  
  # SLIDE ROLLING AVERAGES FOR DIFFERENT WEEKLY PERIODS
  tk_augment_slidify(
    n_desem_lag28,
    .f = ~mean(., na.rm = TRUE),
    .period = c(7, 14, 28, 28*2),
    .align = "center",
    .partial = TRUE
  ) %>%
  ungroup()

data_prepared_tbl <- full_data_tbl %>% 
  filter(!is.na(n_desem),
         !is.na(n_desem_lag28))

future_data_tbl <- full_data_tbl %>% 
  filter(is.na(n_desem))
```

Aquí va el TIME SERIES FEATURE ENGINEERING!!!! definiendo el objeto full_data_table

# Split 28 days

```{r}
splits <- data_prepared_tbl %>% 
  time_series_split(fecha_desembolso, assess = 28, cumulative = TRUE)

splits %>% 
  tk_time_series_cv_plan() %>% 
  plot_time_series_cv_plan(fecha_desembolso, n_desem)
```

# Recipe

```{r}
recipe_spec <- recipe(n_desem ~ ., data = training(splits)) %>% 
  update_role(fecha_desembolso, new_role = "id") %>% 
  step_timeseries_signature(fecha_desembolso) %>% 
  step_rm(matches("(.xts$)|(.iso$)|(hour)|(minute)|(second)|(am.pm)")) %>% 
  step_dummy(all_nominal(), one_hot = TRUE)

# Before the recipe
recipe_spec %>% summary()
# After the recipe
recipe_spec %>% prep() %>% summary()

prep(recipe_spec) %>% juice() %>% glimpse()
```

Need to add lags for this to work!!

# Models

## Model 1: XGBoost

```{r}
wf_fit_xgb_default <- workflow() %>% 
  add_model(boost_tree(mode = "regression") %>%
            set_engine("xgboost")) %>% 
  add_recipe(recipe_spec) %>% 
  fit(training(splits))
```

## Model 6: some other

```{r}

```

# Model table

```{r}
models_tbl <- modeltime_table(
  wf_fit_xgb_default
)

models_tbl
```

# Calibrate

```{r}
calibration_tbl <- models_tbl %>%
    modeltime_calibrate(new_data = testing(splits))
```

# Performance

```{r}
calibration_tbl %>%
    modeltime_accuracy() %>%
    table_modeltime_accuracy()
```

# Forecast

```{r}
test_forecast_tbl <- calibration_tbl %>%
    modeltime_forecast(
        new_data    = testing(splits),
        actual_data = data_prepared_tbl,
        keep_data = TRUE
    ) 

test_forecast_tbl %>% 
  plot_modeltime_forecast()
```

# Visualization filters

```{r}
zona_items <- c("1", "2", "3", "4", "desconocida")

# Por zona
test_forecast_tbl %>% 
  filter(identifier %in% zona_items) %>%
  group_by(identifier) %>% 
  filter_by_time(.start_date = last(fecha_desembolso) %-time% "6 month",
                 .end_date = "end") %>% 
  plot_modeltime_forecast(
    .facet_ncol = 2
  )

# Por tipo de vehículo
test_forecast_tbl %>%
  filter(category %in% "estado_id") %>%
  group_by(identifier) %>% 
  filter_by_time(.start_date = last(fecha_desembolso) %-time% "6 month",
                 .end_date = "end") %>% 
  plot_modeltime_forecast()

# Todas las zonas
test_forecast_tbl %>%
  filter(category %in% "all_zones_id") %>%
  filter_by_time(.start_date = last(fecha_desembolso) %-time% "6 month",
                 .end_date = "end") %>% 
  plot_modeltime_forecast()
```

# Accuracy by zona_id and estado_id

```{r}
accuracy_by_identifier_tbl <- test_forecast_tbl %>% 
  select(category, identifier, .model_desc, .index, .value) %>% 
  pivot_wider(
    names_from = .model_desc,
    values_from = .value
  ) %>% 
  filter(!is.na(XGBOOST)) %>% 
  pivot_longer(cols = XGBOOST) %>% 
  group_by(category, identifier, name) %>% 
  summarize_accuracy_metrics(
    truth = ACTUAL,
    estimate = value,
    metric_set = default_forecast_accuracy_metric_set()
  )

## Select the best rmse
best_rmse_by_identifier_tbl <- accuracy_by_identifier_tbl %>% 
  group_by(identifier) %>% 
  slice_min(rmse, n = 1) %>% 
  ungroup()
```

# Refit the best

```{r}
refit_tbl <- calibration_tbl %>%
    modeltime_refit(data = data_prepared_tbl)

future_forecast_tbl <- refit_tbl %>%
    modeltime_forecast(new_data = future_data_tbl, 
                       actual_data = data_prepared_tbl,
                       keep_data = TRUE) 
# Por Zona
future_forecast_tbl %>%
  filter(identifier %in% zona_items) %>%
  group_by(identifier) %>% 
  filter_by_time(.start_date = last(fecha_desembolso) %-time% "6 month",
                 .end_date = "end") %>% 
  plot_modeltime_forecast(
    .facet_ncol = 3
  )

# Por tipo de vehículo
future_forecast_tbl %>%
  filter(category %in% "estado_id") %>%
  group_by(identifier) %>% 
  filter_by_time(.start_date = last(fecha_desembolso) %-time% "6 month",
                 .end_date = "end") %>% 
  plot_modeltime_forecast(
    .facet_ncol = 3
  )

# Todas las zonas
future_forecast_tbl %>%
  filter(category %in% "all_zones_id") %>%
  filter_by_time(.start_date = last(fecha_desembolso) %-time% "6 month",
                 .end_date = "end") %>% 
  plot_modeltime_forecast()
```

# Writing

```{r}
future_forecast_tbl %>%
  filter(.model_desc != "ACTUAL") %>%
  select(-.key, -c(fecha_desembolso:n_desem_lag28_roll_56)) %>%
  write_csv2("../outputs/prueba_3_model_time_serio.csv")
```

